Quasicharacters and the function-sheaf dictionary

### 2014-09-06 crelle-dictionary.tex
Changes sections heading.

%% \tableofcontents %% Just for papers exceeding 50 pages.

@@

\subsection{The category of quasicharacter sheaves}\label{ssec:category}

@@

\begin{proposition} \label{prop:ses_desc}
Consider the short exact sequence in Corollary~\ref{cor:spectral_ses}.
\begin{enumerate}
\item Every class $[\alpha\oplus\beta\oplus\gamma] \in \Hh^2(E^\bullet_G)$ is cohomologous to one with $\gamma=0$.
\item The map $\Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1({\bar G},\EEx))$ is given by $[\alpha\oplus\beta\oplus 0] \mapsto [\beta]$.
\item Suppose $a \in Z^2({\bar G}, \EEx)$ represents a class in $\Hh^2({\bar G},\EEx)$ fixed by Frobenius.
The map $\Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \to \Hh^2(E^\bullet_G)$ is given by $[a] \mapsto [a \oplus 0 \oplus 0]$.
\end{enumerate}
\todo{add more connecting prose}
\end{proposition}
\begin{proof}
Since $\Hh^2(\Weil{}, C^0({\bar G}, \EEx)) = 0$, we may find a $\gamma_1 \in C^1(\Weil{}, C^0({\bar G}, \EEx))$ with $d_{\Weil{}}\gamma_1 = \gamma$.
Subtracting $d \gamma_1$ from $\alpha\oplus\beta\oplus\gamma$, we may assume that $\gamma = 0$.

@@

\subsection{Objects, when $G$ is etale}\label{ssec:SandT}

@@

\subsection{Morphisms}\label{ssec:mor-etale}

@@

\section{Quasicharacter sheaves defined by discrete isogenies} \label{sec:disc-isog}

@@

\subsection{Quasicharacter sheaves on connected commutative algebraic groups}\label{ssec:connected}

@@

\subsection{Quasicharacters for $p$-adic tori} \label{ssec:qc_tori}

As we explained in the Introduction, our original motivation for this paper was to find a geometrization and categorification of quasicharacters of $p$-adic tori. 
 \todo{Is this the right place to recall/define the term quasicharacter?}
This is is now provided by specializing Theorem~\ref{thm:QCXK} to the case when $X_K= T_K\in \mathcal{N}_K$ is an algebraic torus over $K$. 
Let $T$ be a Neron model for $T_K$.
By \cite{bitan:discriminant}*{Eq 3.1}, the special fibre of the component group scheme for $T$ is given by
\[
 \pi_0(T)_{\bFq} = X_*(T_K)_{\mathcal{I}_K},
\]
where $X_*(T_K)$ is the cocharacter lattice of $T_K$ and $\mathcal{I}_K$ is the inertia group for $K$.
Accordingly, by Theorem~\ref{thm:QCXK}, isomorphism classes in $\QC(T)$   becomes
\begin{equation}
0 \to \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{}} \to \QCiso{T_K} \to \Hom_\text{qc}(T(K),\EEx) \to 0.
\end{equation}
Again by Theorem~\ref{thm:QCXK}, the automorphism groups in $\QC(T_K)$ are given, for every quasicharacter sheaf $\qcs{F}$ for $T_K$, by
\begin{equation}
\Aut(\qcs{F}) \iso \Hom(X_*(T_K)_{\Weil{K}},\EEx).
\end{equation}
\todo{This is dual to Lemma 2.6, and isn't true in general.  Is the fact that $X_*(T_K)$ is torsion-free sufficient?} where $\Weil{K}$ is the Weil group for $K$.

@@

\subsection{Weil restriction and quasicharacters}\label{ssec:wrK}

@@

\begin{proof}
Adapting the argument in \cite{bosch-lutkebohmert-reynaud:NeronModels}*{\S 7.6, Prop 6} to include locally finite type N\'eron models,
one may show that $\Res_{R'/R}(X')$ is a N\'eron model for $\Res_{K'/K}(X_{K'})$.
Thus, by the N\'eron mapping property, the canonical closed immersion
\[
X_K\hookrightarrow \Res_{K'/K}(X_{K'})
\]
 extends uniquely to a morphism
\begin{equation}\label{me}
X\to \Res_{R'/R}(X')
\end{equation}
 of smooth $R$-group schemes.
%
Applying the functor $\Gr^R_{n}$ to \eqref{me}
and using \cite{bertrapelle-gonzales:Greenberg}*{Thm 1.1} defines the morphism of smooth group schemes
\begin{equation}\label{men}
f_n: \Gr_{n-1}^R(X) \to \Res_{k'/k} \Gr_{en-1}^{R'}(X'),
\end{equation}
where $e$ is the ramification index of $K'/K$.
Using Lemma~\ref{lem:pullback}, \eqref{men} induces a functor 
\begin{equation}\label{men-pulled}
f_n^* : \QC(\Res_{k'/k} \Gr_{en-1}^{R'}(X'))\to \QC(\Gr_{n-1}^{R}(X)).
\end{equation}
Since 
\[
\left(\Res_{k'/k} \Gr_{en-1}^{R'}(X') \right)(\Fq) = \left(\Gr_{en-1}^{R'}(X')\right)(k'),
\]
it follows from Lemma~\ref{lem:pullback} that the pullback functor \eqref{men-pulled} actually induces
\[
\Hom(X'(R'_{en-1}),\EEx)  \to  \Hom(X(R_{n-1}),\EEx) .
\]
\[ 
\Hom_{en-1}(X'(R'),\EEx) \to \Hom_{n-1}(X(R)),\EEx)
\]
Since $X'$ is a N\'eron model for $X_{K'}$ and $X$ is a N\'eron model for $X_K$,
 this may be re-written as
\[ 
\Hom_{en-1}(X_{K'}(K'),\EEx)= \Hom_{en-1}(X_{K}(K'),\EEx) \to \Hom_{n-1}(X_K(K)),\EEx)
\]
Passing to limits now defines
\[ 
\Hom_\text{qc}(X_K(K'),\EEx) \to \Hom_\text{qc}(X_K(K)),\EEx)
\]
Argue as in Proposition~\ref{prop:csbe} to see that this is indeed restriction of characters.
\end{proof}

>>>>

 \tableofcontents %% Just for papers exceeding 50 pages.

@@

\subsection{Quasicharacter sheaves}\label{ssec:category}

@@

\begin{proposition} \label{prop:ses_desc}
Consider the short exact sequence in Corollary~\ref{cor:spectral_ses}.
\begin{enumerate}
\item Every class $[\alpha\oplus\beta\oplus\gamma] \in \Hh^2(E^\bullet_G)$ is cohomologous to one with $\gamma=0$.
\item The map $\Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1({\bar G},\EEx))$ is given by $[\alpha\oplus\beta\oplus 0] \mapsto [\beta]$.
\item Suppose $a \in Z^2({\bar G}, \EEx)$ represents a class in $\Hh^2({\bar G},\EEx)$ fixed by Frobenius.
The map $\Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \to \Hh^2(E^\bullet_G)$ is given by $[a] \mapsto [a \oplus 0 \oplus 0]$.
\end{enumerate}
\end{proposition}
\begin{proof}
Since $\Hh^2(\Weil{}, C^0({\bar G}, \EEx)) = 0$, we may find a $\gamma_1 \in C^1(\Weil{}, C^0({\bar G}, \EEx))$ with $d_{\Weil{}}\gamma_1 = \gamma$.
Subtracting $d \gamma_1$ from $\alpha\oplus\beta\oplus\gamma$, we may assume that $\gamma = 0$.

@@

\subsection{Objects in the \'etale case}\label{ssec:SandT}

@@

\subsection{Morphisms in the \'etale case}\label{ssec:mor-etale}

@@

\section{Quasicharacter sheaves on connected commutative algebraic groups} \label{sec:disc-isog}

@@

\subsection{The classical function--sheaf dictionary revisited}\label{ssec:connected}

@@

\subsection{Quasicharacter sheaves for $p$-adic tori} \label{ssec:qc_tori}

As we explained in the Introduction, our original motivation for this paper was to find a geometrization and categorification of quasicharacters of $p$-adic tori. 
This is is now provided by specializing Theorem~\ref{thm:QCXK} to the case when $X_K= T_K\in \mathcal{N}_K$ is an algebraic torus over $K$. 
Let $T$ be a Neron model for $T_K$.
By \cite{bitan:discriminant}*{Eq 3.1}, the special fibre of the component group scheme for $T$ is given by
\[
 \pi_0(T)_{\bFq} = X_*(T_K)_{\mathcal{I}_K},
\]
where $X_*(T_K)$ is the cocharacter lattice of $T_K$ and $\mathcal{I}_K$ is the inertia group for $K$.
Accordingly, by Theorem~\ref{thm:QCXK}, isomorphism classes in $\QC(T)$   becomes
\begin{equation}
0 \to \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{}} \to \QCiso{T_K} \to \Hom_\text{qc}(T(K),\EEx) \to 0.
\end{equation}
Again by Theorem~\ref{thm:QCXK}, the automorphism groups in $\QC(T_K)$ are given, for every quasicharacter sheaf $\qcs{F}$ for $T_K$, by
\begin{equation}
\Aut(\qcs{F}) \iso \Hom(X_*(T_K)_{\Weil{K}},\EEx).
\end{equation}
%\todo{This is dual to Lemma 2.6, and isn't true in general.  Is the fact that $X_*(T_K)$ is torsion-free sufficient?} where $\Weil{K}$ is the Weil group for $K$.

@@

\subsection{Weil restriction and quasicharacter sheaves}\label{ssec:wrK}

@@

\begin{proof}
Adapting the argument in \cite{bosch-lutkebohmert-reynaud:NeronModels}*{\S 7.6, Prop 6} to include locally finite type N\'eron models,
one may show that $\Res_{R'/R}(X')$ is a N\'eron model for $\Res_{K'/K}(X_{K'})$.
Thus, by the N\'eron mapping property, the canonical closed immersion
\[
X_K\hookrightarrow \Res_{K'/K}(X_{K'})
\]
 extends uniquely to a morphism
\begin{equation}\label{me}
X\to \Res_{R'/R}(X')
\end{equation}
 of smooth $R$-group schemes.
%
Applying the functor $\Gr^R_{n}$ to \eqref{me}
and using \cite{bertrapelle-gonzales:Greenberg}*{Thm 1.1} defines the morphism of smooth group schemes
\begin{equation}\label{men}
f_n: \Gr_{n-1}^R(X) \to \Res_{k'/k} \Gr_{en-1}^{R'}(X'),
\end{equation}
where $e$ is the ramification index of $K'/K$.
Using Lemma~\ref{lem:pullback}, \eqref{men} induces a functor 
\begin{equation}\label{men-pulled}
f_n^* : \QC(\Res_{k'/k} \Gr_{en-1}^{R'}(X'))\to \QC(\Gr_{n-1}^{R}(X)).
\end{equation}
Since 
\[
\left(\Res_{k'/k} \Gr_{en-1}^{R'}(X') \right)(\Fq) = \left(\Gr_{en-1}^{R'}(X')\right)(k'),
\]
it follows from Lemma~\ref{lem:pullback} that the pullback functor \eqref{men-pulled} actually induces
%\[
%\Hom(X'(R'_{en-1}),\EEx)  \to  \Hom(X(R_{n-1}),\EEx) .
%\]
\[ 
\Hom_{en-1}(X'(R'),\EEx) \to \Hom_{n-1}(X(R)),\EEx)
\]
Since $X'$ is a N\'eron model for $X_{K'}$ and $X$ is a N\'eron model for $X_K$,
 this may be re-written as \todo{I'm going to think about this a bit more.}
\[ 
\Hom_{en-1}(X_{K'}(K'),\EEx)= \Hom_{en-1}(X_{K}(K'),\EEx) \to \Hom_{n-1}(X_K(K)),\EEx)
\]
Passing to limits now defines
\[ 
\Hom_\text{qc}(X_K(K'),\EEx) \to \Hom_\text{qc}(X_K(K)),\EEx)
\]
Argue as in Proposition~\ref{prop:csbe} to see that this is indeed restriction of characters.
\end{proof}
